Joint training of classification model and similarity model for low-resource text classification in chatbot

### 2020-07-25 chatbot_classification/aaai_2021/paper.tex
modified:   paper.pdf   modified:   paper.tex
update model structure figures and Abstract and Introduction Section

%% LyX 2.3.1-1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass{article}
\usepackage[latin9]{inputenc}
\usepackage{array}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{graphicx}

@@

\begin{document}
\title{     
  SFC: Few-Shot Text Classification via Similarity Fused with Classification System
}
\name{Jingwen Huang}   
\address{     
  \{hanmei613\}@gmail.com   
}
\maketitle

\begin{abstract}
  Building  conversational chatbot system has become a popular solution to sharing
  the  work  of customer service under various business scenarios. A conversational chatbot needs to detect user's intent given a few words, which essentially equals to short-text classification problem
  in  the  field  of  Natural  Language  Processing.  However, each time for a new
  service,  the  task-specific chatbot system often needs to perform well in few-shot setups due
  to  lack  of  domain-specific  data,  which  is  still quite hard even if we use
  powerful  pretrained  model  like  Roberta. Therefore, in this paper, we propose
  SFC,  a  system  fusion  of  both  similarity  model and classification model to
  overcome  this  challenge.  Our main contributions are: 1) transfer learning and
  negative  sampling  based  on  sentence-pair  are utilized as remedy for lack of
  data;  2)  multi-task  learning is involved to achieve faster training speed and
  better  performance;  3) model ensembling of classification and similarity model
  guarantees  inference  speed  while keeping high accuracy. Additionally, we also
  conduct  extensive  experiments on four public datasets in few-shot setup (i.e.,
  with  only  5 to 20 training data per class). The experimental results show that
  our system can steadily outperform several competitive baselines by 2 percent in
  average accuracy.
\end{abstract}
%\keywords{}

\section{Introduction}
\label{sec:intro}

Single-turn  conversational  chatbots  are  designed to transform existing tasks
that  rely  on human agents, such as classifying customers' questions or queries
to  find  the  corresponding  answer,  into an automatic process based on intent
classification   model.  Since  user  queries  are  usually  much  shorter  than
paragraphs  or  documents,  the chatbot can actually be turned into a short-text
Classification\cite{phan2008learning,yan2009dynamic,hua2015short} task  in  Natural  Language Processing. Moreover, at the initial
stage  of  building  chatbots, it's usually extremely hard to collect sufficient
data  for each class. In this way, building a single-turn conversational chatbot
become  a  short-text  classification\cite{sriram2010short} problem under few-shot setting\cite{yu2018diverse}.

Short  texts\cite{song2014short} are usually more ambiguous in comparison with long texts since they
don't contain enough contextual information, which poses a great challenge for
classification\cite{chen2019deep}. In addition, the few-shot scenario\cite{yu2018diverse} adds  even more difficulties to the classification task since there is no
enough  information  for  the  model  to  learn  for  each  class.  Comparing to
non-pre-trained neural network such as convolutional neural networks (CNNs)\cite{kim2014convolutional,zhang2015character,conneau2016very} and long short term memory networks (LSTMs)\cite{mousa2017contextual,liu2016recurrent}, the
recently introduced pre-trained language models on large corpus like BERT\cite{devlin2018bert}  and  RoBERTa\cite{liu2019roberta} has been proven more powerful in
solving  many  NLP  tasks  including  short-text  classification  for  deficient
data\cite{madabushi2020cost}.  Especially  for  few-shot  scenarios, transfer
learning  based  on  pre-trained  model  tends  to  do more help to the negative
effects brought by scarcity of training data. Due to the fact that RoBERTa is an
improved  version  of  BERT,  we  build  our SFC chatbot system using RoBERTa as
pre-trained  context-dependent  embeddings.  To  our  knowledge, the most common
approach  is  adding  a  softmax  classifier to the top of RoBERTa(i.e., RoBERTa
classifier),  which  turns  the  problem into a simple text classification task.
However,  despite  the  fact  that  RoBERTa has achieved amazing results in many
Natural  Language  Processing tasks, we still believe it has much more potential
under few-shot setting.

Therefore,  in  this  paper  we  propose  to  involve  similarity  model  (i.e.,
sentence-pair  classifier)  based  on  RoBERTa,  which  can  score  the semantic
similarity between two sentences by multiple cross-attention mechanism\cite{barkan2020scalable}.  A natural idea to enhance model performance in few-shot setting is
to  further  pre-train RoBERTa with target domain data for transfer learning\cite{sun2019fine}.  However, it's always not easy to find domain-specific data for
certain  service/product  if  we  try  to further pre-train a RoBERTa classifier
directly.  In  comparison,  it's  relatively  feasible  to  obtain  dataset  for
semantically duplicate sentence pair identification task. That is to say, we can
obtain a further pre-trained similarity model to identify the class with highest
semantic  similarity  level to each user query. Afterwards, we can fine-tune the
similarity  model using sentence pairs sampled from target task dataset based on
negative  sampling  strategy\cite{bamler2020extreme}, which can provide us with
more features to learn from limited amount of data.

We  also apply multi-task learning\cite{caruana1993multitask,collobert2008unified} in
training  process  to  enhance the training speed and performance. We set up two
different  objectives  for  training. The first one is the regular sentence-pair
similarity  score  which  help the model learn the semantic similarity between a
query  to each existing data point. The second target is to learn the similarity
score  between  a  query  and all the data points of a certain class as a whole.
Specific  knowledge contained in these two tasks can be fully explored to obtain
a faster training speed and higher accuracy.

Another  contribution of this work is that we ensemble the similarity model with
classification  model  at  the  inference  step.  In  contrast with common model
ensembling  methods\cite{breiman1996stacked,schapire1998boosting},  the classification model are used as an auxiliary model in
our  system  to  select promising sentence-pair candidates for similarity model.
The  experiment  results  on  4  short-text  classification  datasets in various
few-shot  settings  show that our system can outperform single model baseline by
at  least  2 percent on average. Besides, we can also control the inference time
for one single query to be within 0.5 seconds, which makes the system applicable
in real-life single-turn chatbot scenario.

>>>>

%% LyX 2.3.1-1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass{article}
\usepackage[latin9]{inputenc}
\usepackage{array}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{float}
\usepackage{subfigure}

@@

\begin{document}
\title{     
  %SFC: Few-Shot Text Classification via Similarity Fused with Classification System
  Joint training of classification model and similarity model for low-resource
  text classification in chatbot
}
\name{Jingwen Huang}   
\address{     
  \{hanmei613\}@gmail.com   
}
\maketitle

\begin{abstract}
  Building  conversational  chatbot  system  has  become  a  popular solution to customer  services  under various business scenarios. A
  conversational  chatbot needs to detect user's intent given a few words, which
  essentially  equals  to  short-text  classification  problem  in  the field of Natural  Language  Processing.  Moreover,  each  time  for  a  new service, the task-specific  chatbot  system  often needs to perform well in few-shot setups due  to lack of domain-specific data, which is quite a challenge for single-model system such as text classification model system or sentence-pair semantic similarity model system under such a low-resource condition. Therefore, in this paper, we propose
  SFC,  a 2-stage joint system with multi-task training technique for both text classification task and sentence-pair semantic similarity task to
  overcome  this challenge. Furthermore, we additionally propose an improved version of SFC to allow text classification model and sentence-pair model be combined into a joint model organized in hierarchical structure. We also
  conduct extensive experiments on 4 public and 1 private datasets in few-shot setup (i.e.,
  with only 5 to 20 training data per class). The experimental results show that
  our  system can steadily outperform several competitive single-model baselines by 2 percent
  in average accuracy.
\end{abstract}
%\keywords{}

\section{Introduction}
\label{sec:intro}
% latest version start
Task-specific conversational chatbot\cite{wen2016network} are designed to share the working pressure of human customer service agents who are responsible for solving customers' questions or queries about certain products or services. Regardless of whether the conversation is single-turn or multi-turn, the essential technical solution behind the chatbot is intent classification model which can identify the correct intent behind user's input text and find the corresponding answer.

Although task-specific conversational chatbot already has a wide application in various business and industries, it's still a quite challenging task due to its natural properties of low-resource. First, customers' utterance within a conversation is usually quite short and completed in a few words. Short text\cite{song2014short} are generally more ambiguous in comparison with long texts such as paragraphs or documents since they don't contain enough contextual information, which poses a great challenge\cite{chen2019deep} for classification task\cite{phan2008learning,yan2009dynamic,hua2015short}. Second, at the initial stage of building a chatbot for specific task or service, it's often extremely hard to collect sufficient data samples for each class. The reason is that it's usually too expensive and time consuming to extract different ways of natural language expression for each intent class from history conversation log or even manually compose the corpus from nothing. Therefore, the key to building a task-specific chatbot with high performance becomes solving the challenge of short-text classification\cite{sriram2010short} problem under few-shot setting\cite{yu2018diverse}.

\begin{figure}[htbp]
\centering 
\subfigure[text classification model structure based on pre-trained transformer]{
\label{Fig.1.sub.1}
\includegraphics[scale=0.6]{picture1.pdf}}
\subfigure[sentence-pair model structure based on cross-attention mechanism]{
\label{Fig.1.sub.2}
\includegraphics[scale=0.55]{picture2.pdf}}
\caption{Model Structure of 2 popular approach for building task-specific conversational chatbot}
\label{Fig.1.main}
\end{figure}

One of the most popular approach among existing work was text classification model based on neural network. Since the length of the text is quite short, many previous work\cite{wen2016network} choose neural network such as convolutional neural networks (CNNs)\cite{kim2014convolutional,zhang2015character,conneau2016very}
or long short term memory networks
(LSTMs)\cite{mousa2017contextual,liu2016recurrent} to accomplish the task of extracting semantic feature from limited amount of words. A common model structure is adding a softmax classifier to the top of the neural network. Afterwards, pre-trained language models on large corpus like BERT\cite{devlin2018bert} and
RoBERTa\cite{liu2019roberta} has  been proven more powerful in solving many NLP
tasks      including      short-text      classification\cite{madabushi2020cost}.   Especially   for  few-shot  scenarios\cite{yu2018diverse}, pre-trained  model  based on transformers\cite{vaswani2017attention}, shown in Fig. \ref{Fig.1.sub.1}, tends  to  do more help to reducing the negative effects brought by scarcity of training data.

Another popular approach among previous work was based on sentence-pair model. The motivation of this approach was started from the idea of Information Retrival (IR) based chatbot\cite{jafarpour2010filter, leuski2011npceditor}. Having a Q-A (Question-Answer) pairs dataset and user query Q, the IR based conversational system will look up in the Q-A dataset for the pair (Q', A') that best matches query Q through semantic analysis and returns A' as the answer to Q\cite{mnasri2019recent}. In this way, sentence-pair model pre-trained on large corpus of semantic similarity identification task can be applied for being used as a tool to identify the class with highest semantic similarity to customer's query. A common model structure of pre-trained\cite{devlin2018bert} sentence-pair model is based on multiple cross-attention mechanism\cite{barkan2020scalable}, shown in Fig. \ref{Fig.1.sub.2}. Due to the fact that many experiment results have shown that RoBERTa\cite{liu2019roberta} is an improved  version  of  BERT, and has achieved amazing results on both text classification and sentence-pairs semantic similarity tasks, we choose RoBERTa as an important baseline approach in this paper.

Despite the success of these 2 approaches, they still have some limitations in task-specific chatbot scenario. As for the text classification model approach, it's quite hard to use data augmentation method to solve the data insufficiency challenge since the it's usually unfeasible to get large amount of domain-specific data when facing a new task. With respect to the sentence-pair model approach, though we can obtain large amount of data in semantically duplicate sentence pair identification domain for transfer learning\cite{sun2019fine}, it's still quite hard to perform well on task-specific dataset because the training objective of sentence-pair model is semantic similarity, which is slightly different from the target of classification task. That is to say, there always exists some intent classes which cannot be distinguished by each other merely depending on semantic similarity. For example, we have 2 user query saying, A: What should I do if I want to change my password? B: What is the modification rule if I want to change my password? In this case, sentence A is semantically similar to B, since they both express the 
desire for changing password. However, it's still reasonable to classify them into 2 intent classes, since A is asking for the method for changing password, while B is asking for the rule to follow when creating a new password.

\begin{figure}[htbp]
\begin{centering}
\includegraphics[scale=0.48]{picture3}
\par\end{centering}
\caption{\textbf{Network Structure of 2-stage SFC:} a joint system of multi-task for both text classification task and sentence-pair semantic similarity task.}
\label{fig:ver.1.3}
\end{figure}

The above limitation motivates us to propose a joint system of both text classification task and sentence-pair semantic similarity task, named SFC here, shown in Fig. \ref{fig:ver.1.3}. Our goal is to utilize the advantage of the feasibility of transfer learning based on semantic similarity, while in the mean time, add the classification task's target into the training process of sentence-pair model for multi-task learning\cite{caruana1993multitask,collobert2008unified}. To obtain such a joint system, we first start with preparation work, which is pre-training a sentence-pair model on external corpus for sentence pair duplication identification, and then find an auxiliary model or tool which can help us sample out top-K most related intent classes. This auxiliary model or tool can either be searching engine such as elasticsearch\cite{divya2013elasticsearch} or a text classification model trained on task-specific chatbot data. Afterwards, we further fine-tune the sentence-pair model with 2 training tasks: 1) the regular sentence pair similarity tasks. Here we use the text classification model as an auxiliary model in this paper to sample sentence pairs for training based on negative sampling strategy\cite{bamler2020extreme}. The reason is that a task-specific chatbot often has hundreds of intent classes, which forms too many sentence pairs for training since any two sentences from two different classes can form a sentence-pair of negative sample. Besides, according to Bamler\cite{bamler2020extreme}, training on negative samples from most confusing wrong label can help model converge faster and obtain better performance. 2) the classification task on top-K (here K is a hyperparameter) candidate classes provided by our auxiliary model or tool. Here we use the average pooling of the representation given by sentence-pair model for sentence pairs formed by the user input query and the corpus sentences in each class as the feature. We stack the task-specific layers on the top of the shared-parameter sentence-pair model structure, which is a classic parameter sharing mechanism\cite{collobert2008unified, subramanian2018learning,liu2019multi} for multi-task\cite{sun2019fine}. In this way, specific knowledge contained in these 2 tasks can be fully explored and deeply interact with each other to obtain better overall performance.

However, since the auxiliary tool in joint system mentioned above works in a separate stage from sentence-pair model, we find the the quality of sampled candidate classes might limit the overall performance of SFC, since the candidate classes for each user query is always fixed during the multi-task training process. This observation motivates us to further improve SFC into a joint model of sentence-pair model and text classification model in a hierarchical architecture by putting different tasks at different network layers\cite{sogaard2016deep, hashimoto2016joint} and then training them together, shown in Fig \ref{fig:framework}. Here, the text classification model involved in joint training process replace the role of auxiliary tools. In this way, the sentence pairs selected from top-K classes becomes dynamic, which means the text classification model will also be further optimized to provide better top-K classes pooling result.

We summarize our contributions as follow:

1) We propose a novel joint system named SFC with multi-task training technique designed for task-specific conversational chatbot with low-resource, in which we make full utilization of the advantages brought by both the text classification task and sentence-pair semantic similarity task.

2) To further improve the system performance of SFC, we also propose an innovative joint model structure, in which the text classification model and sentence-pair model are fused into one single model for joint training.

3) Experiment results on 4 public and 1 private short-text classification datasets with respect to intent recognition tasks under conversational chatbot scenario all demonstrated that our proposed SFC joint system with multi-task, as well as the improved version of SFC, can both achieve remarkable improvement comparing to some powerful single-task and single-model baselines, especially under few-shot settings.

\begin{figure*}[t]
\begin{centering}
\includegraphics[scale=0.72]{picture4} 
\par\end{centering}
\caption{\textbf{Network Structure of hierarchical SFC:} a joint model of sentence-pair model and text classification model organized in hierarchical structure by completing different tasks at different network layers.}
\label{fig:framework}
\end{figure*}

% latest version end
